---
title: "Clinical Capsule Pipeline — Extended Pilot Study Example"
format:
  html:
    toc: true
    toc-depth: 3
vignette: >
  %\VignetteIndexEntry{Clinical Capsule Pipeline — Extended Pilot Study Example}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

## What this article is for

This article is a **worked, end-to-end walkthrough** of the clinical capsule
pipeline on the **bundled CDISC Pilot Study** outputs shipped with ksAI
(`inst/examples/pilot-study/meta`).

It is intentionally long. Each section explains not only *what* to call, but
*why* that step exists, how capsules are shaped for that table type, and what
you should inspect before moving on. Use it as:

- a companion to the [Architecture and Pipeline](Architecture_and_Pipeline.html)
  article (concepts → practice);
- a template for wiring your own study through
  `as_capsules()` → `ks_annotate()` → `ks_embed()` → `ks_retrieve()` → `ks_reason()`;
- a reference for LM Studio setup when running the live LLM / embedding steps.

Live model calls require a local LM Studio server. Structure, inventory, and
keyword annotation steps run without any model. Set `RUN_LLM <- FALSE` below to
stop before network-backed steps.

---

## Why a capsule pipeline (instead of dumping full tables)

Medical writers rarely need every row of every table for one question. Full
injection of a large AE or laboratory table into an 8k-token local model
fails or forces aggressive `max_rows` truncation that silently drops evidence.

The capsule pipeline solves that by changing **what** the reasoning model sees:

| Approach | What the model receives | Best for |
|----------|-------------------------|----------|
| Direct `ks_llm(study, ids=...)` | Entire selected table(s) as markdown/compact | 1–2 known, modest tables |
| Capsule pipeline | Only top-*n* concept summaries (and optional children) | Many tables, exploratory Q&A, token limits |

A **clinical capsule** is a concept-centric unit (for example one SOC, one PT,
one demographic parameter) with:

- **traceability** — `source_id` + `source_rows` point back to the table;
- **compact evidence** — `compact_text` and parsed `stats` for the LLM;
- **hierarchy** — `parent_id` / `child_ids` for progressive disclosure;
- **retrieval handles** — `domain`, `keywords`, optional `embedding`.

```{mermaid}
flowchart LR
  subgraph prep["Build once"]
    L["ks_load()"]
    C["as_capsules()"]
    A["ks_annotate()"]
    E["ks_embed()"]
    S["save_capsules()"]
  end
  subgraph ask["Ask many times"]
    R["ks_retrieve()"]
    Q["ks_reason()"]
  end
  L --> C --> A --> E --> S
  S --> R --> Q
```

Build the store once per study (or per writing campaign). Reuse the `.ksc`
file across sessions without reloading ksTFL JSON.

---

## Prerequisites

### Bundled pilot data

ksAI ships a self-contained meta folder from the CDISC Pilot Study
(ksTFL `save_report()` artefacts):

```
inst/examples/pilot-study/meta/
  _index.json
  *.json          # specs + data
  *.svg           # figures (ignored by capsule build for tables)
```

Resolve it at runtime with:

```{r}
#| eval: false
meta_path <- system.file(
  "examples", "pilot-study", "meta",
  package = "ksAI"
)
dir.exists(meta_path)
```

In a development checkout before install, use:

```{r}
#| eval: false
meta_path <- file.path("inst", "examples", "pilot-study", "meta")
```

### Models (optional live steps)

This walkthrough is written for **LM Studio** with three roles:

| Role | Example model | Used by |
|------|---------------|---------|
| Domain + semantic (small) | `qwen3.5-4b` | `as_capsules(..., model = ...)`, `ks_annotate(..., model = ...)` |
| Reasoning / CSR (large) | `qwen3.6-14b-a3b-fablevibes` | `ks_reason()` |
| Reasoning / CSR (large) | `gemma-4-14b-a4b-heretic-v2` | `ks_reason()` |
| Embeddings | `text-embedding-nomic-embed-text-v1.5` | `ks_embed()`, query vectors in `ks_retrieve()` |

**URL rule that matters:**

- Chat / annotate / reason: `http://127.0.0.1:1234` (**no** `/v1`)
- Embeddings: `http://127.0.0.1:1234/v1` (**with** `/v1`)

ellmer's `chat_lmstudio()` appends `/v1` itself. Passing `/v1` into chat
constructors makes model discovery fail even when the model is loaded.

### Packages and options

```{r}
library(ksAI)

# Capsule demos keep tables modest so compact_text stays readable.
# The full pilot AE table can exceed 200 capsules at max_rows = 200;
# 20–50 is enough to demonstrate hierarchy and retrieval.
ks_set_option(
  max_rows = 40L,
  context_format = "compact",
  provider = "lm_studio",
  embed_model = "text-embedding-nomic-embed-text-v1.5",
  embed_url = "http://127.0.0.1:1234/v1"
)

RUN_LLM <- FALSE   # TRUE when LM Studio is up with the models above
CHAT_URL <- "http://127.0.0.1:1234"
MODEL_SMALL <- "qwen3.5-4b"
MODEL_LARGE <- "qwen3.6-14b-a3b-fablevibes"
#MODEL_LARGE <- "gemma-4-14b-a4b-heretic-v2"

OUT_DIR <- file.path(tempdir(), "ksAI-capsule-pilot")
dir.create(OUT_DIR, recursive = TRUE, showWarnings = FALSE)
```

---

## Step 0 — Choose outputs that span clinical domains

Domain tags on capsules are inferred by `.capsule_infer_domain()` in a
**language-agnostic** order:

1. Explicit `annotations$domain` via `enrich_context()`
2. Session `domain_map` (`ks_set_option()`) — exact output id or regex
3. Structural cues (`ROW_KIND` SOC/PT/HLGT/… → `AE`)
4. Optional small LLM when `as_capsules(..., model = ..., llm_domain = "always")`
5. Multilingual title/subtitle/source lexicon (EN, DE, FR, ES, IT, RU, ZH, JA, …)
6. ICH/CSR-style numbering in the output id (`14-5.x` → AE, `14-6.x` → LB, …)
7. Optional small LLM when still `UNKNOWN` (`model` + default `llm_domain = "unknown"`)
8. `"UNKNOWN"`

Bare English `"baseline"` alone is **not** mapped to `DM` (avoids tagging
efficacy change-from-baseline tables as demographics). Prefer
`enrich_context(..., annotations = list(domain = ...))`, `domain_map`, or a
small local model for study-specific / non-English titles:

```{r}
#| eval: false
store <- as_capsules(
  study,
  model = "qwen3.5-4b",
  provider = "lm_studio",
  base_url = "http://127.0.0.1:1234",
  llm_domain = "unknown"   # or "always" after hard signals
)
# Later: re-tag remaining UNKNOWN domains during annotate
store <- ks_annotate(store, model = "qwen3.5-4b", provider = "lm_studio",
                     base_url = "http://127.0.0.1:1234")
```

| Domain | Typical cues (any language) | Typical pilot tables |
|--------|-----------------------------|----------------------|
| `AE` | adverse / НДЯ / 不良事件; MedDRA SOC/PT | `14-5.01`, `14-5.02` |
| `DM` | demograph / демограф / 人口学 | `14-2.01` |
| `VS` | vital / витальн / バイタル | `14-7.01`, `14-7.02` |
| `LB` | lab / лаборатор / 实验室 | `14-6.01`, `14-6.02`, … |
| `EFFC` | efficacy / эффективност / 疗效 | `14-3.01`, `14-3.02`, … |
| `EX` / `DS` | exposure / disposition (lexicon or id) | — |
| `UNKNOWN` | no match (until LLM / override) | figures, untitled stubs |

We deliberately load **six tables across five domains** so retrieval filters
and domain inventory are non-trivial:

```{r}
meta_path <- system.file("examples", "pilot-study", "meta", package = "ksAI")

TABLES <- c(
  demographics   = "14-2.01",  # DM  — Age, Sex, Race, …
  efficacy_adas  = "14-3.01",  # EFFC
  efficacy_cibic = "14-3.02",  # EFFC
  adverse_events = "14-5.01",  # AE   — SOC / PT hierarchy
  laboratory     = "14-6.01",  # LB   — visit / parameter rows
  vital_signs    = "14-7.02"   # VS   — change from baseline
)

# Discover without loading data JSON:
catalog <- ks_list_ids(meta_path)
catalog[catalog$id %in% TABLES, c("id", "type", "title")]
```

---

## Step 1 — Load a targeted study

```{r}
study <- ks_load(meta_path, ids = unname(TABLES))
study
# 6 tables, 0 figures, 0 texts (for this subset)
```

Each `study[["14-5.01"]]` is a `ks_context`: title, population, columns, span
headers, rendered rows (with `section` / `kind`), footnotes. Capsules are
built from those rows — they never re-open the original ksTFL session.

Inspect one context before capsule construction:

```{r}
ctx_ae <- study[["14-5.01"]]
ctx_ae
cat(substr(as_compact(ctx_ae), 1, 700), "\n...\n")
```

You should see the compact DSL: `SPANS:` legend, SOC section headers, PT lines
with arm-keyed `n(%)` values. That same DSL is what each capsule stores in
`compact_text` for its row group.

---

## Step 2 — Structure agent: `as_capsules()`

### What the structure agent does

`as_capsules(study)` walks every loaded context and:

1. Infers a **domain** once per output (rules; optional `model` /
   `llm_domain` / `llm_min_confidence` when a small LLM is available).
2. Identifies a **row-label column** (first visible column).
3. Assigns each row (or section block) a **level**:
   - AE-style: `OVERALL` → `SOC` → `PT` (from `ROW_KIND` / section headers);
   - continuous / demo / lab: `OVERALL` → `PARAM`.
4. Builds synthetic **SOC** capsules when a section header exists without an
   explicit SOC kind row.
5. Parses arm-wise **stats** where possible.
6. Renders **compact_text** for that row slice only.
7. Wires **parent_id** / **child_ids**.
8. Links **linked_ids** to text listings in the same inferred domain (if any
   text outputs were loaded).

```{r}
store <- as_capsules(study)
# Optional domain LLM (same signature as Step 0):
# store <- as_capsules(study, model = MODEL_SMALL, provider = PROVIDER,
#                      base_url = CHAT_URL, llm_domain = "unknown")
store
length(store$capsules)
```

### Inventory by source table and domain

Always print an inventory before annotation. It surfaces domain mis-tags and
unexpected capsule counts early.

```{r}
inventory <- do.call(rbind, lapply(names(store$capsules), function(cid) {
  cap <- store$capsules[[cid]]
  data.frame(
    capsule_id = cid,
    source_id  = cap$source_id,
    domain     = cap$domain,
    level      = cap$level,
    label      = cap$label,
    n_rows     = length(cap$source_rows),
    n_children = length(cap$child_ids),
    stringsAsFactors = FALSE
  )
}))

# Counts by table × domain × level
print(as.data.frame(table(
  source = inventory$source_id,
  domain = inventory$domain
)))

print(as.data.frame(table(
  source = inventory$source_id,
  level  = inventory$level
)))
```

### Anatomy of one capsule (inspect deeply)

Pick one AE SOC capsule and one demographic PARAM capsule:

```{r}
# Exact IDs depend on labels; search rather than hard-coding when exploring:
cardiac_ids <- grep("CARDIAC", names(store$capsules), value = TRUE)
cardiac_ids

cap_soc <- store$capsules[[cardiac_ids[[1]]]]
cap_soc

# Traceability — which table rows produced this capsule?
cap_soc$source_id
cap_soc$source_rows

# Hierarchy
cap_soc$parent_id
cap_soc$child_ids   # PT capsules under this SOC

# Evidence the LLM will see later
cat(cap_soc$compact_text)

# Parsed statistics (list structure; arm / measure dependent)
str(cap_soc$stats, max.level = 2)

# Semantic slots (empty until annotate / embed)
cap_soc$concepts
cap_soc$keywords
cap_soc$embedding   # NULL until ks_embed()
```

**What “good” looks like for an AE SOC capsule:**

- `level == "SOC"`, `domain == "AE"`;
- `child_ids` lists PT capsule ids under that SOC;
- `compact_text` contains arm-keyed incidence lines, not the whole AE table;
- `source_rows` is a contiguous or section-scoped index set you could audit
  against `study[["14-5.01"]]$rows`.

Compare with a demographics capsule:

```{r}
age_ids <- grep("AGE|Age", names(store$capsules), value = TRUE, ignore.case = TRUE)
cap_age <- store$capsules[[age_ids[[1]]]]
cap_age$level      # typically PARAM
cap_age$domain     # typically DM
cat(cap_age$compact_text)
```

PARAM capsules are flatter: fewer children, one characteristic or visit row,
still arm-spanned compact text when the source table uses span headers.

### Persist the empty (structure-only) store

Even before enrichment, save the store so you can resume without rebuilding:

```{r}
path_struct <- file.path(OUT_DIR, "pilot_capsules_structure.ksc")
save_capsules(store, path_struct)
# Later: store <- load_capsules(path_struct)
```

---

## Step 3 — Semantic agent: `ks_annotate()`

### Pass 1 — deterministic keywords (no model)

`ks_annotate(store)` without a `model` argument always runs a pure-R pass:

- tokenizes `label` + `compact_text` (+ existing concepts);
- drops a small stopword list;
- maps known abbreviations to synonyms (`teae` → “treatment emergent adverse
  event”, `soc` → “system organ class”, …);
- seeds `concepts` from the label when empty.

```{r}
store <- ks_annotate(store)

# Keywords are now retrieval fodder even without embeddings
cap_soc <- store$capsules[[cardiac_ids[[1]]]]
head(cap_soc$keywords, 30)
cap_soc$synonyms
cap_soc$concepts
```

**Why this pass matters:** keyword overlap is 30% of the default
`ks_retrieve()` score. A capsule with empty keywords and no embedding ranks
only on metadata filters. Always run at least the R pass before retrieval
demos.

### Pass 2 — optional small-LLM enrichment

A 4B local model can add clinical concepts and synonyms that pure tokenization
misses (for example mapping “SINUS BRADYCARDIA” to broader cardiac terms).
The same `model` argument also reclassifies capsules still tagged `UNKNOWN`
(once per `source_id`). Use `force_domain = TRUE` to reclassify every table.
This is slower and memory-heavy: unload the 26B model in LM Studio first if
needed.

```{r}
if (RUN_LLM) {
  # Annotate a representative subset first — full stores can be large.
  demo_ids <- unique(c(
    cardiac_ids[1],
    age_ids[1],
    grep("14-3.01", names(store$capsules), value = TRUE)[1],
    grep("14-6.01", names(store$capsules), value = TRUE)[1],
    grep("14-7.02", names(store$capsules), value = TRUE)[1]
  ))

  demo <- store
  demo$capsules <- store$capsules[demo_ids]

  demo <- ks_annotate(
    demo,
    model              = MODEL_SMALL,
    provider           = "lm_studio",
    base_url           = CHAT_URL,
    force              = TRUE,   # re-run even if concepts already seeded
    force_domain       = FALSE,  # only repair UNKNOWN domains
    llm_min_confidence = 0.5
  )

  store$capsules[demo_ids] <- demo$capsules

  for (cid in demo_ids) {
    cat("\n== ", cid, " ==\n", sep = "")
    print(store$capsules[[cid]]$concepts)
    print(store$capsules[[cid]]$synonyms)
  }
}
```

The LLM is instructed to return JSON with keys `concepts`, `synonyms`,
`keywords`. Non-JSON wrappers are stripped when possible; failed parses leave
the capsule unchanged for that field.

---

## Step 4 — Embeddings: `ks_embed()`

Embeddings turn capsule `compact_text` into dense vectors for semantic
similarity. Without them, `ks_retrieve()` still works (keyword + metadata
only); with them, natural-language queries align better to clinical phrasing
that does not share surface tokens with the table.

```{r}
if (RUN_LLM) {
  store <- ks_embed(store)   # uses embed_model / embed_url options

  n_emb <- sum(vapply(
    store$capsules,
    function(c) !is.null(c$embedding),
    logical(1)
  ))
  cat("Embedded", n_emb, "of", length(store$capsules), "capsules\n")

  # Dimensionality depends on the embedding model (often 768 for nomic):
  length(store$capsules[[cardiac_ids[[1]]]]$embedding)
}
```

**Operational notes:**

- Capsules with empty `compact_text` may be skipped on embed failure.
- Re-running `ks_embed()` skips capsules that already have vectors unless
  `force = TRUE`.
- Persist after embedding so you do not re-hit the endpoint every session:

```{r}
path_full <- file.path(OUT_DIR, "pilot_capsules.ksc")
save_capsules(store, path_full)
```

---

## Step 5 — Retrieval agent: `ks_retrieve()`

### Scoring model

For each capsule, `ks_retrieve()` computes:

$$
\text{score} = w_s \cdot \text{semantic} + w_k \cdot \text{keyword} + w_m \cdot \text{metadata}
$$

Defaults: $w_s = 0.6$, $w_k = 0.3$, $w_m = 0.1$.

| Signal | Range | Meaning |
|--------|-------|---------|
| Semantic | 0–1 (cosine) | Query embedding vs `capsule$embedding`; 0 if either missing |
| Keyword | 0–1 | Fraction of query tokens found in keywords ∪ label |
| Metadata | 0–1 | Fraction of supplied `filter` fields that match exactly |

Filters (`domain`, `level`, `population`) **do not hard-exclude** capsules in
the current implementation; they contribute to the metadata score. Prefer
post-filtering the subset or inspecting scores when you need a hard domain
gate.

### Domain-specific queries on the pilot store

```{r}
queries <- list(
  ae_cardiac = list(
    query  = "cardiac disorders and bradycardia in the high dose arm",
    filter = list(domain = "AE", level = "SOC")
  ),
  efficacy = list(
    query  = "ADAS-Cog change from baseline week 24 treatment difference",
    filter = list(level = "PARAM")
  ),
  laboratory = list(
    query  = "laboratory summary statistics change from baseline over visits",
    filter = list(domain = "LB")
  ),
  vital_signs = list(
    query  = "systolic blood pressure change from baseline end of treatment",
    filter = list(level = "PARAM")
  ),
  demographics = list(
    query  = "age sex race baseline demographic characteristics",
    filter = list(domain = "DM", level = "PARAM")
  )
)

if (RUN_LLM) {
  subsets <- lapply(names(queries), function(nm) {
    q <- queries[[nm]]
    cat("\n######## ", nm, " ########\n", sep = "")
    sub <- ks_retrieve(
      store,
      query  = q$query,
      n      = 5L,
      filter = q$filter
    )
    print(sub)
    print(sub$scores)
    sub
  })
  names(subsets) <- names(queries)
} else {
  # Offline smoke: keyword-only retrieval still ranks something useful
  sub_offline <- ks_retrieve(
    store,
    query = "cardiac bradycardia",
    n = 5L,
    filter = list(domain = "AE"),
    # Without embeddings, semantic weight is wasted — shift mass to keywords:
    weights = list(semantic = 0, keyword = 0.8, metadata = 0.2)
  )
  print(sub_offline)
  print(sub_offline$scores)
}
```

**How to read `scores`:**

- High `semantic`, low `keyword` — paraphrased clinical language matched the
  embedding space.
- High `keyword`, low `semantic` — shared tokens (or no embeddings).
- High `metadata` — filter fields matched; still compare against raw rank.

Inspect the winning capsule’s evidence before reasoning:

```{r}
# Using the offline or first live subset:
top_id <- names(sub_offline$capsules)[[1]]
cat(store$capsules[[top_id]]$compact_text)
```

---

## Step 6 — Reasoning agent: `ks_reason()`

`ks_reason()` is retrieve + generate in one call:

1. `ks_retrieve(store, query, n)` (same scoring as above);
2. optional **progressive disclosure** — if `expand = TRUE`, pull every
   `child_ids` capsule from the store into the prompt;
3. format capsule metadata + `compact_text` into a context block;
4. chat with a larger model under a system prompt that forbids inventing
   statistics.

```{r}
if (RUN_LLM) {
  # 6a. Safety narrative at SOC level, then expand PT children
  out_ae <- ks_reason(
    store,
    query    = "Summarize cardiac disorder incidence across treatment arms. Cite specific n(%) values from the capsules.",
    n        = 6L,
    expand   = TRUE,
    model    = MODEL_LARGE,
    provider = "lm_studio",
    base_url = CHAT_URL
  )
  out_ae
  cat(out_ae$response)

  # 6b. Efficacy question — usually PARAM capsules, expand often unnecessary
  out_eff <- ks_reason(
    store,
    query    = "What treatment differences for ADAS-Cog change from baseline at Week 24 appear in the capsules?",
    n        = 6L,
    expand   = FALSE,
    model    = MODEL_LARGE,
    provider = "lm_studio",
    base_url = CHAT_URL
  )
  cat(out_eff$response)

  # 6c. Cross-domain question — do not over-filter; let retrieval pick
  out_mix <- ks_reason(
    store,
    query    = "Are there safety signals that should contextualize interpretation of the primary efficacy results?",
    n        = 5L,
    expand   = FALSE,
    model    = MODEL_LARGE,
    provider = "lm_studio",
    base_url = CHAT_URL
  )
  cat(out_mix$response)

  save_result(out_ae, file.path(OUT_DIR, "reason-cardiac"))
  save_result(out_eff, file.path(OUT_DIR, "reason-adas"))
}
```

### Progressive disclosure in practice

```
expand = FALSE  →  only the top-n retrieved capsules
expand = TRUE   →  top-n plus all of their child_ids (e.g. SOC → PTs)
```

Use `expand = TRUE` when the question needs preferred-term detail under a
system organ class. Use `FALSE` when SOC-level (or PARAM-level) summaries
already answer the question — smaller prompts, less noise, lower chance of
context overflow on local 8k models.

### Guardrails you should still enforce as a writer

The model is instructed not to invent numbers, but you must:

1. Check that cited `n(%)` values appear in the printed capsule
   `compact_text` or original table;
2. Remember `max_rows` truncated the source table — rare PTs below the row
   cap never became capsules;
3. Treat domain tags as retrieval aids, not regulatory classifications.

---

## Step 7 — Chain capsule reasoning into CSR drafting

Capsule answers are often drafts. A common pattern is:

1. `ks_reason()` over retrieved capsules → safety/efficacy bullets;
2. `save_result()`;
3. `ks_llm()` on a **small** related table with `prior =` that result to draft
   a CSR paragraph, verifying numbers against the table context.

Keep the second step on a compact table (ADAS or CIBIC, not the full AE
listing). Prior text plus a large table overflows local contexts.

```{r}
if (RUN_LLM) {
  prior <- load_result(file.path(OUT_DIR, "reason-cardiac"))

  out_csr <- ks_llm(
    study,
    ids      = TABLES[["efficacy_cibic"]],
    skill    = "csr_section",
    title    = "Efficacy in the context of cardiac safety findings",
    prompt   = paste(
      "Using the prior capsule-based cardiac analysis as background,",
      "draft a short CSR note on how CIBIC+ results should be read",
      "alongside the cardiac safety signal.",
      "Do not invent AE counts; refer to the prior for safety numbers.",
      "Verify any efficacy numbers against the table context."
    ),
    prior    = prior,
    model    = MODEL_LARGE,
    provider = "lm_studio",
    base_url = CHAT_URL
  )

  save_result(out_csr, file.path(OUT_DIR, "csr-efficacy-with-safety-prior"))
  cat(out_csr$response)
}
```

---

## Step 8 — Operational checklist for your own study

1. **Load targeted ids** — start with the domains your question needs; add
   tables later rather than always loading the full meta folder.
2. **Set `max_rows` deliberately** — capsules cannot recover rows never
   imported.
3. **`as_capsules(study)`** — inventory `domain` × `level` × `source_id`;
   optional `model` / `llm_domain` for UNKNOWN titles.
4. **Fix or accept domain tags** — use
   `enrich_context(..., annotations = list(domain = ...))`,
   `domain_map`, or `ks_annotate(..., model = ..., force_domain = ...)` for
   non-English / custom titles.
5. **`ks_annotate(store)`** — always; add `model =` for concepts/keywords and
   UNKNOWN domain repair on high-value capsules if needed.
6. **`ks_embed(store)`** — when an embedding endpoint is available.
7. **`save_capsules()`** — `.ksc` is the reusable artefact.
8. **`ks_retrieve()` / `ks_reason()`** — tune `n`, `expand`, `filter`, and
   `weights` per question type.
9. **Audit** — every claim against `compact_text` or the source
   `ks_context`.
10. **Chain** — optional `ks_llm(..., prior =)` on a small table for CSR
    polish.

---

## Runnable script twin

The same pilot meta folder powers the executable script:

```r
# Development checkout
Rscript inst/examples/pilot-study-workflows.R

# Installed package
Rscript system.file(
  "examples", "pilot-study-workflows.R",
  package = "ksAI"
)
```

That script covers direct LLM skills, fact filtering, capsules, and chaining
in one file. This article isolates the **capsule pipeline** and documents the
decisions behind each call.

---

## Design takeaways

- Capsules are the **unit of retrieval and reasoning**, not whole tables.
- Structure is deterministic R; semantics can be cheap (keywords) or richer
  (small LLM + embeddings).
- Reasoning models should see **evidence subsets**, with PT detail pulled on
  demand via `expand`.
- Traceability (`source_id`, `source_rows`) is what makes capsule-based CSR
  drafts auditable in a regulated setting.
- The bundled pilot study is enough to exercise AE hierarchy, lab/visit
  PARAMs, demographics, vitals, and efficacy tables in one store — use the
  inventory tables in Step 2 as your ground truth before trusting retrieval
  filters.
