---
title: "Clinical Capsule Pipeline — Extended Pilot Study Example"
format:
  html:
    toc: true
    toc-depth: 3
vignette: >
  %\VignetteIndexEntry{Clinical Capsule Pipeline — Extended Pilot Study Example}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

## What this article is for

This article is a **worked, end-to-end walkthrough** of the clinical capsule
pipeline on the **bundled CDISC Pilot Study** outputs shipped with ksAI
(`inst/examples/pilot-study/meta`).

It is intentionally long. Each section explains not only *what* to call, but
*why* that step exists, how capsules are shaped by the LLM, and what you
should inspect before moving on. Use it as:

- a companion to the [Architecture and Pipeline](Architecture_and_Pipeline.html)
  article (concepts → practice);
- a template for wiring your own study through
  `as_capsules()` → `ks_annotate()` → `ks_embed()` → `ks_retrieve()` → `ks_reason()`;
- a reference for LM Studio setup when running the live LLM / embedding steps.

Capsule **formation** (`as_capsules()`) always needs a chat model — there is no
rule-based / CDISC domain path. Keyword annotation can still run offline once
you have a store. Set `RUN_LLM <- FALSE` below to skip the optional LLM
annotate pass, embeddings, live retrieval scoring, and reasoning (you can still
inspect a previously saved `.ksc`).

---

## Why a capsule pipeline (instead of dumping full tables)

Medical writers rarely need every row of every table for one question. Full
injection of a large AE or laboratory table into an 8k-token local model
fails or forces aggressive `max_rows` truncation that silently drops evidence.

The capsule pipeline solves that by changing **what** the reasoning model sees:

| Approach | What the model receives | Best for |
|----------|-------------------------|----------|
| Direct `ks_llm(study, ids=...)` | Entire selected table(s) as markdown/compact | 1–2 known, modest tables |
| Capsule pipeline | Only top-*n* concept summaries (and optional children) | Many tables/figures, exploratory Q&A, token limits |

A **clinical capsule** is a named semantic unit (for example “cardiac adverse
events”, “ADAS-Cog efficacy”, “vital signs change from baseline”) with:

- **membership** — `member_ids` lists whole table/figure output ids (one output
  may belong to several capsules);
- **compact evidence** — `compact_text` concatenates member excerpts for the LLM;
- **hierarchy** — `parent_id` / `child_ids` from the LLM tree (not MedDRA
  `ROW_KIND`);
- **retrieval handles** — `label`, `keywords`, optional `embedding`, and
  optional `population`.

```{mermaid}
flowchart LR
  subgraph prep["Build once"]
    L["ks_load()"]
    C["as_capsules()"]
    A["ks_annotate()"]
    E["ks_embed()"]
    S["save_capsules()"]
  end
  subgraph ask["Ask many times"]
    R["ks_retrieve()"]
    Q["ks_reason()"]
  end
  L --> C --> A --> E --> S
  S --> R --> Q
```

Build the store once per study (or per writing campaign). Reuse the `.ksc`
file across sessions without reloading ksTFL JSON.

---

## Prerequisites

### Bundled pilot data

ksAI ships a self-contained meta folder from the CDISC Pilot Study
(ksTFL `save_report()` artefacts):

```
inst/examples/pilot-study/meta/
  _index.json
  *.json          # specs + data
  *.svg           # figures (included when loaded; vision models get pixels)
```

Resolve it at runtime with:

```{r}
#| eval: false
meta_path <- system.file(
  "examples", "pilot-study", "meta",
  package = "ksAI"
)
dir.exists(meta_path)
```

In a development checkout before install, use:

```{r}
#| eval: false
meta_path <- file.path("inst", "examples", "pilot-study", "meta")
```

### Models (optional live steps)

This walkthrough is written for **LM Studio** with three roles:

| Role | Example model | Used by |
|------|---------------|---------|
| Capsule formation + semantic (small) | `qwen3.5-4b` | `as_capsules(..., model = ...)`, `ks_annotate(..., model = ...)`, optional `ks_review_capsules()` |
| Reasoning / CSR (large) | `qwen3.6-14b-a3b-fablevibes` | `ks_reason()` |
| Reasoning / CSR (large) | `gemma-4-14b-a4b-heretic-v2` | `ks_reason()` |
| Embeddings | `text-embedding-nomic-embed-text-v1.5` | `ks_embed()`, query vectors in `ks_retrieve()` |

Prefer a **vision-capable** small model when the study includes figures —
`as_capsules(..., attach_images = TRUE)` attaches readable image assets so the
classifier can use plot pixels, not only titles.

**URL rule that matters:**

- Chat / annotate / reason: `http://127.0.0.1:1234` (**no** `/v1`)
- Embeddings: `http://127.0.0.1:1234/v1` (**with** `/v1`)

ellmer's `chat_lmstudio()` appends `/v1` itself. Passing `/v1` into chat
constructors makes model discovery fail even when the model is loaded.

### Packages and options

```{r}
library(ksAI)

# Capsule demos keep tables modest so compact excerpts stay readable.
# The full pilot AE table can be large at max_rows = 200;
# 20–50 is enough to demonstrate hierarchy and retrieval.
PROVIDER <- "lm_studio"
ks_set_option(
  max_rows = 40L,
  context_format = "compact",
  provider = PROVIDER,
  embed_model = "text-embedding-nomic-embed-text-v1.5",
  embed_url = "http://127.0.0.1:1234/v1"
)

RUN_LLM <- FALSE   # TRUE when LM Studio is up with the models above
CHAT_URL <- "http://127.0.0.1:1234"
MODEL_SMALL <- "qwen3.5-4b"
MODEL_LARGE <- "qwen3.6-14b-a3b-fablevibes"
#MODEL_LARGE <- "gemma-4-14b-a4b-heretic-v2"

OUT_DIR <- file.path(tempdir(), "ksAI-capsule-pilot")
dir.create(OUT_DIR, recursive = TRUE, showWarnings = FALSE)
```

---

## Step 0 — Choose outputs that span clinical topics

`as_capsules()` does **not** assign CDISC domains. The LLM groups loaded
**tables and figures** by meaning into a named capsule tree. Membership is by
whole output id (`member_ids`); multi-membership is allowed (one AE table can
sit under both a safety root and a cardiac child).

We deliberately load **six tables across five clinical topics** so retrieval
and inventory are non-trivial:

| Topic | Typical pilot tables |
|-------|----------------------|
| Demographics | `14-2.01` |
| Efficacy (ADAS / CIBIC) | `14-3.01`, `14-3.02` |
| Adverse events | `14-5.01` |
| Laboratory | `14-6.01` |
| Vital signs | `14-7.02` |

```{r}
meta_path <- system.file("examples", "pilot-study", "meta", package = "ksAI")

TABLES <- c(
  demographics   = "14-2.01",
  efficacy_adas  = "14-3.01",
  efficacy_cibic = "14-3.02",
  adverse_events = "14-5.01",
  laboratory     = "14-6.01",
  vital_signs    = "14-7.02"
)

# Discover without loading data JSON:
catalog <- ks_list_ids(meta_path)
catalog[catalog$id %in% TABLES, c("id", "type", "title")]
```

Figures in the same meta folder can be loaded alongside tables. When
`attach_images = TRUE` (default), readable `asset_path` images are sent to
vision-capable models during classification and deep review.

---

## Step 1 — Load a targeted study

```{r}
study <- ks_load(meta_path, ids = unname(TABLES))
study
# 6 tables, 0 figures, 0 texts (for this subset)
```

Each `study[["14-5.01"]]` is a `ks_context`: title, population, columns, span
headers, rendered rows (with `section` / `kind`), footnotes. Capsule
`compact_text` is built from member excerpts of these contexts — formation
never re-opens the original ksTFL session.

Inspect one context before capsule construction:

```{r}
ctx_ae <- study[["14-5.01"]]
ctx_ae
cat(substr(as_compact(ctx_ae), 1, 700), "\n...\n")
```

You should see the compact DSL: `SPANS:` legend, SOC section headers, PT lines
with arm-keyed `n(%)` values. Similar excerpts (truncated by
`max_excerpt_rows`) are what the structure LLM sees in the classify catalog.

---

## Step 2 — Structure agent: `as_capsules()`

### What the structure agent does

`as_capsules(study, model = ...)` is **LLM-only**. It:

1. Builds a catalog of Table/Figure contexts (titles, populations, compact or
   full excerpts, figure assets when available).
2. Asks the chat model to propose a named semantic tree (`capsule_id`,
   `label`, `member_ids`, `parent_id`, confidence).
3. Batches large catalogs (`batch_size`) and merges partial trees when needed.
4. Drops capsules below `min_confidence`.
5. Materializes each capsule with `compact_text` from its members and wires
   `parent_id` / `child_ids`.

There is **no** rule-based domain inference, no MedDRA `ROW_KIND` hierarchy,
and no `linked_ids` / CDISC domain chain. Hierarchy is whatever meaning the
LLM assigns.

`model` is required. **`n_ctx` cannot be set in `as_capsules()`** — raise
Context Length when loading the model in LM Studio / Ollama.

Two practical call patterns:

**Option A — small context** (e.g. `n_ctx = 8192`): shrink each classify
prompt. Partial trees are merged by a second LLM pass. Good when you cannot
raise context; weaker for subtle cross-output multi-membership than a single
full-catalog call.

```{r}
#| eval: false
store <- as_capsules(
  study,
  model = MODEL_SMALL,
  provider = PROVIDER,
  base_url = CHAT_URL,
  detail = "compact",
  max_excerpt_rows = 4L,
  batch_size = 1L,
  attach_images = FALSE,
  params = ellmer::params(temperature = 0),
  api_args = list(enable_thinking = FALSE)
)
```

**Option B — large context** (raise Context Length in LM Studio, e.g. 32k+):
larger batches / excerpts; better whole-catalog interpretation. Use
`attach_images = TRUE` only with a vision-capable model.

```{r}
#| eval: false
store <- as_capsules(
  study,
  model = MODEL_SMALL,
  provider = PROVIDER,
  base_url = CHAT_URL,
  detail = "compact",
  max_excerpt_rows = 12L,
  batch_size = 12L,
  attach_images = FALSE,  # TRUE only if the model has vision
  params = ellmer::params(temperature = 0),
  api_args = list(enable_thinking = FALSE)
)
```

This walkthrough uses Option A by default so it fits typical local 8k windows:

```{r}
store <- as_capsules(
  study,
  model = MODEL_SMALL,
  provider = PROVIDER,
  base_url = CHAT_URL,
  detail = "compact",
  max_excerpt_rows = 4L,
  batch_size = 1L,
  attach_images = FALSE,
  params = ellmer::params(temperature = 0),
  api_args = list(enable_thinking = FALSE)
)
store
length(store$capsules)
```

### Inventory by label and members

Always print an inventory before annotation. It surfaces odd labels, empty
membership, and unexpected multi-membership early.

```{r}
inventory <- do.call(rbind, lapply(names(store$capsules), function(cid) {
  cap <- store$capsules[[cid]]
  data.frame(
    capsule_id = cid,
    label      = cap$label,
    members    = paste(cap$member_ids, collapse = ", "),
    n_members  = length(cap$member_ids),
    parent_id  = cap$parent_id %||% NA_character_,
    n_children = length(cap$child_ids),
    stringsAsFactors = FALSE
  )
}))
inventory

# Which capsules claim each output?
for (cid in names(store$capsules)) {
  cap <- store$capsules[[cid]]
  cat(cid, ": ", cap$label, " — members ",
      paste(cap$member_ids, collapse = ", "), "\n", sep = "")
}
```

Structural review helpers (no LLM for `review_capsules` / `capsule_tree`):

```{r}
print(review_capsules(store, study))   # empty / unknown members / cycles / orphans
capsule_tree(store)                    # printable parent → child tree

# Membership table (output_id × capsule); multi-membership shows n_capsules > 1
print(capsule_membership(store, study))
```

Optional deep critique with a (vision-capable) model:

```{r}
if (RUN_LLM) {
  rev <- ks_review_capsules(
    store,
    study,
    model = MODEL_SMALL,
    provider = PROVIDER,
    base_url = CHAT_URL
  )
  cat(rev$response)
}
```

Inspect one capsule’s full member renders from the live study (not the
truncated build excerpt):

```{r}
sample_id <- names(store$capsules)[[1]]
cat(capsule_content(store, sample_id, study, format = "compact"))
```

### Anatomy of one capsule (inspect deeply)

Pick capsules by **label** (exact ids are LLM-assigned slugs):

```{r}
# Exact IDs depend on the model; search labels rather than hard-coding:
cardiac_ids <- names(store$capsules)[
  grepl("cardiac|bradycardia|heart",
        vapply(store$capsules, function(c) c$label, character(1)),
        ignore.case = TRUE)
]
cardiac_ids

cap <- store$capsules[[cardiac_ids[[1]]]]
cap

# Membership — which outputs belong to this concept?
cap$member_ids

# Hierarchy (LLM tree, not MedDRA ROW_KIND)
cap$parent_id
cap$child_ids

# Evidence the reasoning model will see later
cat(cap$compact_text)

# Semantic slots (empty until annotate / embed)
cap$concepts
cap$keywords
cap$embedding   # NULL until ks_embed()
```

**What “good” looks like:**

- `label` is a readable clinical concept name (any language the titles use);
- `member_ids` are real study output ids (audit with `review_capsules()`);
- multi-membership is coherent (same table under related parent/child labels);
- `compact_text` is a member excerpt bundle, not the entire study;
- `child_ids` refine the parent when the tree has depth.

Compare with a demographics- or efficacy-oriented capsule:

```{r}
age_ids <- names(store$capsules)[
  grepl("age|demographic|baseline characteristic",
        vapply(store$capsules, function(c) c$label, character(1)),
        ignore.case = TRUE)
]
if (length(age_ids)) {
  cap_age <- store$capsules[[age_ids[[1]]]]
  cap_age$label
  cap_age$member_ids
  cat(cap_age$compact_text)
}
```

### Persist the store after formation

Save early so you can resume without re-calling the structure model:

```{r}
path_struct <- file.path(OUT_DIR, "pilot_capsules_structure.ksc")
save_capsules(store, path_struct)
# Later: store <- load_capsules(path_struct)
```

---

## Step 3 — Semantic agent: `ks_annotate()`

### Pass 1 — deterministic keywords (no model)

`ks_annotate(store)` without a `model` argument always runs a pure-R pass:

- tokenizes `label` + `compact_text` (+ existing concepts);
- drops a small stopword list;
- maps known abbreviations to synonyms (`teae` → “treatment emergent adverse
  event”, `soc` → “system organ class”, …);
- seeds `concepts` from the label when empty.

```{r}
store <- ks_annotate(store)

# Keywords are now retrieval fodder even without embeddings
cap <- store$capsules[[cardiac_ids[[1]]]]
head(cap$keywords, 30)
cap$synonyms
cap$concepts
```

**Why this pass matters:** keyword overlap is 30% of the default
`ks_retrieve()` score. A capsule with empty keywords and no embedding ranks
only on metadata filters. Always run at least the R pass before retrieval
demos.

### Pass 2 — optional small-LLM enrichment

A 4B local model can add clinical concepts and synonyms that pure tokenization
misses (for example mapping “SINUS BRADYCARDIA” to broader cardiac terms).
`model` is optional and only used for concept / synonym / keyword extraction.

```{r}
if (RUN_LLM) {
  # Annotate a representative subset first — full stores can be large.
  demo_ids <- unique(c(
    head(cardiac_ids, 1L),
    head(age_ids, 1L),
    head(names(store$capsules), 3L)
  ))

  demo <- store
  demo$capsules <- store$capsules[demo_ids]

  demo <- ks_annotate(
    demo,
    model    = MODEL_SMALL,
    provider = PROVIDER,
    base_url = CHAT_URL,
    force    = TRUE   # re-run even if concepts already seeded
  )

  store$capsules[demo_ids] <- demo$capsules

  for (cid in demo_ids) {
    cat("\n== ", cid, " ==\n", sep = "")
    print(store$capsules[[cid]]$concepts)
    print(store$capsules[[cid]]$synonyms)
  }
}
```

The LLM is instructed to return JSON with keys `concepts`, `synonyms`,
`keywords`. Non-JSON wrappers are stripped when possible; failed parses leave
the capsule unchanged for that field.

---

## Step 4 — Embeddings: `ks_embed()`

Embeddings turn capsule `compact_text` into dense vectors for semantic
similarity. Without them, `ks_retrieve()` still works (keyword + metadata
only); with them, natural-language queries align better to clinical phrasing
that does not share surface tokens with the table.

```{r}
if (RUN_LLM) {
  store <- ks_embed(store)   # uses embed_model / embed_url options

  n_emb <- sum(vapply(
    store$capsules,
    function(c) !is.null(c$embedding),
    logical(1)
  ))
  cat("Embedded", n_emb, "of", length(store$capsules), "capsules\n")

  # Dimensionality depends on the embedding model (often 768 for nomic):
  length(store$capsules[[cardiac_ids[[1]]]]$embedding)
}
```

**Operational notes:**

- Capsules with empty `compact_text` may be skipped on embed failure.
- Re-running `ks_embed()` skips capsules that already have vectors unless
  `force = TRUE`.
- Persist after embedding so you do not re-hit the endpoint every session:

```{r}
path_full <- file.path(OUT_DIR, "pilot_capsules.ksc")
save_capsules(store, path_full)
```

---

## Step 5 — Retrieval agent: `ks_retrieve()`

### Scoring model

For each capsule, `ks_retrieve()` computes:

$$
\text{score} = w_s \cdot \text{semantic} + w_k \cdot \text{keyword} + w_m \cdot \text{metadata}
$$

Defaults: $w_s = 0.6$, $w_k = 0.3$, $w_m = 0.1$.

| Signal | Range | Meaning |
|--------|-------|---------|
| Semantic | 0–1 (cosine) | Query embedding vs `capsule$embedding`; 0 if either missing |
| Keyword | 0–1 | Fraction of query tokens found in keywords ∪ label |
| Metadata | 0–1 | Fraction of supplied `filter` fields that match |

Filters accept `label`, `member_id`, and/or `population`. They **do not
hard-exclude** capsules in the current implementation; they contribute to the
metadata score. Prefer post-filtering the subset or inspecting scores when you
need a hard gate (for example “must include table `14-5.01`”).

### Content queries on the pilot store

```{r}
queries <- list(
  ae_cardiac = list(
    query  = "cardiac disorders and bradycardia in the high dose arm",
    filter = list(member_id = "14-5.01")
  ),
  efficacy = list(
    query  = "ADAS-Cog change from baseline week 24 treatment difference",
    filter = list(member_id = "14-3.01")
  ),
  laboratory = list(
    query  = "laboratory summary statistics change from baseline over visits",
    filter = list(member_id = "14-6.01")
  ),
  vital_signs = list(
    query  = "systolic blood pressure change from baseline end of treatment",
    filter = list(label = "vital")
  ),
  demographics = list(
    query  = "age sex race baseline demographic characteristics",
    filter = list(member_id = "14-2.01")
  )
)

if (RUN_LLM) {
  subsets <- lapply(names(queries), function(nm) {
    q <- queries[[nm]]
    cat("\n######## ", nm, " ########\n", sep = "")
    sub <- ks_retrieve(
      store,
      query  = q$query,
      n      = 5L,
      filter = q$filter
    )
    print(sub)
    print(sub$scores)
    sub
  })
  names(subsets) <- names(queries)
} else {
  # Offline smoke: keyword-only retrieval still ranks something useful
  sub_offline <- ks_retrieve(
    store,
    query = "cardiac bradycardia",
    n = 5L,
    filter = list(member_id = "14-5.01"),
    # Without embeddings, semantic weight is wasted — shift mass to keywords:
    weights = list(semantic = 0, keyword = 0.8, metadata = 0.2)
  )
  print(sub_offline)
  print(sub_offline$scores)
}
```

**How to read `scores`:**

- High `semantic`, low `keyword` — paraphrased clinical language matched the
  embedding space.
- High `keyword`, low `semantic` — shared tokens (or no embeddings).
- High `metadata` — filter fields matched; still compare against raw rank.

Inspect the winning capsule’s evidence before reasoning:

```{r}
# Using the offline or first live subset:
top_id <- names(sub_offline$capsules)[[1]]
cat(store$capsules[[top_id]]$compact_text)
```

---

## Step 6 — Reasoning agent: `ks_reason()`

`ks_reason()` is retrieve + generate in one call:

1. `ks_retrieve(store, query, n)` (same scoring as above);
2. optional **progressive disclosure** — if `expand = TRUE`, pull every
   `child_ids` capsule from the store into the prompt;
3. format capsule metadata + `compact_text` into a context block;
4. chat with a larger model under a system prompt that forbids inventing
   statistics.

```{r}
if (RUN_LLM) {
  # 6a. Safety narrative, then expand child capsules under retrieved parents
  out_ae <- ks_reason(
    store,
    query    = "Summarize cardiac disorder incidence across treatment arms. Cite specific n(%) values from the capsules.",
    n        = 6L,
    expand   = TRUE,
    model    = MODEL_LARGE,
    provider = PROVIDER,
    base_url = CHAT_URL
  )
  out_ae
  cat(out_ae$response)

  # 6b. Efficacy question — expand often unnecessary if leaves already answer
  out_eff <- ks_reason(
    store,
    query    = "What treatment differences for ADAS-Cog change from baseline at Week 24 appear in the capsules?",
    n        = 6L,
    expand   = FALSE,
    model    = MODEL_LARGE,
    provider = PROVIDER,
    base_url = CHAT_URL
  )
  cat(out_eff$response)

  # 6c. Cross-topic question — do not over-filter; let retrieval pick
  out_mix <- ks_reason(
    store,
    query    = "Are there safety signals that should contextualize interpretation of the primary efficacy results?",
    n        = 5L,
    expand   = FALSE,
    model    = MODEL_LARGE,
    provider = PROVIDER,
    base_url = CHAT_URL
  )
  cat(out_mix$response)

  save_result(out_ae, file.path(OUT_DIR, "reason-cardiac"))
  save_result(out_eff, file.path(OUT_DIR, "reason-adas"))
}
```

### Progressive disclosure in practice

```
expand = FALSE  →  only the top-n retrieved capsules
expand = TRUE   →  top-n plus all of their child_ids
```

Use `expand = TRUE` when the question needs finer detail under a broad parent
capsule (for example a safety root whose children split cardiac vs hepatic).
Use `FALSE` when the retrieved leaves already answer the question — smaller
prompts, less noise, lower chance of context overflow on local 8k models.

### Guardrails you should still enforce as a writer

The model is instructed not to invent numbers, but you must:

1. Check that cited `n(%)` values appear in the printed capsule
   `compact_text`, `capsule_content()`, or original table;
2. Remember `max_rows` / `max_excerpt_rows` truncated what the structure model
   and capsules could see — rare rows below the cap may never appear;
3. Treat capsule labels as retrieval aids, not regulatory classifications.

---

## Step 7 — Chain capsule reasoning into CSR drafting

Capsule answers are often drafts. A common pattern is:

1. `ks_reason()` over retrieved capsules → safety/efficacy bullets;
2. `save_result()`;
3. `ks_llm()` on a **small** related table with `prior =` that result to draft
   a CSR paragraph, verifying numbers against the table context.

Keep the second step on a compact table (ADAS or CIBIC, not the full AE
listing). Prior text plus a large table overflows local contexts.

```{r}
if (RUN_LLM) {
  prior <- load_result(file.path(OUT_DIR, "reason-cardiac"))

  out_csr <- ks_llm(
    study,
    ids      = TABLES[["efficacy_cibic"]],
    skill    = "csr_section",
    title    = "Efficacy in the context of cardiac safety findings",
    prompt   = paste(
      "Using the prior capsule-based cardiac analysis as background,",
      "draft a short CSR note on how CIBIC+ results should be read",
      "alongside the cardiac safety signal.",
      "Do not invent AE counts; refer to the prior for safety numbers.",
      "Verify any efficacy numbers against the table context."
    ),
    prior    = prior,
    model    = MODEL_LARGE,
    provider = PROVIDER,
    base_url = CHAT_URL
  )

  save_result(out_csr, file.path(OUT_DIR, "csr-efficacy-with-safety-prior"))
  cat(out_csr$response)
}
```

---

## Step 8 — Operational checklist for your own study

1. **Load targeted ids** — start with the tables/figures your question needs;
   add outputs later rather than always loading the full meta folder.
2. **Set `max_rows` deliberately** — capsules cannot recover rows never
   imported; `max_excerpt_rows` further limits what the structure LLM sees.
3. **`as_capsules(study, model = ...)`** — required model; inventory labels and
   `member_ids`; review with `review_capsules()`, `capsule_tree()`,
   `capsule_membership()`, and optionally `ks_review_capsules()`.
4. **Accept multi-membership** — one output in several capsules is intentional;
   audit with `capsule_membership()` rather than forcing a single domain tag.
5. **`ks_annotate(store)`** — always; add `model =` for concepts/keywords on
   high-value capsules if needed (no domain re-tagging).
6. **`ks_embed(store)`** — when an embedding endpoint is available.
7. **`save_capsules()`** — `.ksc` is the reusable artefact.
8. **`ks_retrieve()` / `ks_reason()`** — tune `n`, `expand`,
   `filter` (`label` / `member_id` / `population`), and `weights` per question.
9. **Audit** — every claim against `compact_text`, `capsule_content()`, or the
   source `ks_context`.
10. **Chain** — optional `ks_llm(..., prior =)` on a small table for CSR
    polish.

---

## Runnable script twin

The same pilot meta folder powers the executable script:

```r
# Development checkout
Rscript inst/examples/pilot-study-workflows.R

# Installed package
Rscript system.file(
  "examples", "pilot-study-workflows.R",
  package = "ksAI"
)
```

That script covers direct LLM skills, fact filtering, capsules, and chaining
in one file. This article isolates the **capsule pipeline** and documents the
decisions behind each call.

---

## Design takeaways

- Capsules are the **unit of retrieval and reasoning**, not whole tables.
- Formation is **LLM-only**: named semantic tree + `member_ids`, with optional
  figure images for vision models — not CDISC domains or MedDRA `ROW_KIND`.
- Semantics can be cheap (keywords) or richer (small LLM + embeddings).
- Reasoning models should see **evidence subsets**, with child detail pulled on
  demand via `expand`.
- Traceability is via `member_ids` plus `capsule_content()` / source
  `ks_context` — that audit path is what makes capsule-based CSR drafts
  usable in a regulated setting.
- The bundled pilot study is enough to exercise AE, lab/visit, demographics,
  vitals, and efficacy content in one store — use the inventory in Step 2 as
  your ground truth before trusting retrieval filters.
