---
title: "Getting Started with ksAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ksAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

## What ksAI does

`ksAI` adds an offline, AI-native reasoning layer on top of the
[ksTFL](https://crow16384.github.io/ksTFL/) clinical tables framework. ksTFL
computes the tables, figures, and listings of a study deterministically and
writes them to a *meta folder* as JSON (structural metadata plus the rendered
data). `ksAI` reads that folder, compiles every output into a self-contained,
LLM-ready **study registry**, and lets a medical writer ask questions across
the whole set of statistical results and draft Clinical Study Report (CSR)
narratives — using a local model via Ollama or LM Studio, or a hosted
provider.

No live ksTFL session is required: `ksAI` works entirely from the saved JSON.

## 1. Load a study

Point `load_study()` at a ksTFL meta folder (the `metaPath` you passed to
`ksTFL::save_report()`):

```{r}
library(ksAI)

study <- load_study("path/to/outputs/meta")
study
```

Each output becomes a `ks_context`: the render-join of specification and data,
with the title, analysis population, columns, span headers, rendered rows, and
footnotes all in one object.

```{r}
study$tables[["14-3.01"]]
```

Rows keep their displayed values *and* the structural metadata a writer needs —
the section a row belongs to (`Baseline`, `Week 24`, `Change from Baseline`)
and whether it is a detail or a label row.

## 2. Open a chat session

`ks_chat()` wires the study to a model. The context strategy is chosen
automatically from the study size: small studies embed every table directly in
the system prompt; larger studies embed a compact index and register tools so
the model can pull details on demand.

```{r}
chat <- ks_chat(study, model = "qwen3:14b", provider = "ollama")
```

## 3. Ask questions across the whole study

`ask()` is free-form and conversational. This is where cross-table reasoning
happens — relating efficacy, safety, vital signs, and quality-of-life outputs:

```{r}
ask(chat, "How do the vital sign changes relate to the adverse event profile?")
ask(chat, "Which tables support the primary efficacy conclusion, and what do they show?")
ask(chat, "Are the analysis populations consistent across the efficacy tables?")
```

## 4. Draft CSR text with skills

Skills are reusable prompt templates. `ks_llm()` fills a skill's placeholders
from the study context and returns the model's response.

```{r}
# Describe one table
ks_llm(chat, skill = "describe", id = "14-3.01")

# Audience-tailored summary
ks_llm(chat, skill = "summarize", id = "14-3.01", audience = "clinician")

# Draft a CSR results section
ks_llm(chat, skill = "csr_section", id = "14-3.01", title = "ADAS-Cog (11)")

# Narrated consistency review of two tables
compare_tables(chat, "14-3.01", "14-3.02")
```

## 5. Customize skills

The built-in skills are `describe`, `summarize`, `csr_section`, and `review`.
Add your own — or override a built-in — by pointing `skills_dir` at a folder of
Markdown templates. Placeholders use `{{name}}` syntax; `{{context}}` is filled
automatically.

```{r}
ks_set_option(skills_dir = "~/my-study-prompts")
ks_list_skills()

# A file ~/my-study-prompts/exec_summary.md becomes a skill:
ks_llm(chat, skill = "exec_summary", id = "14-3.01")
```

## 6. Persist the study

Compile once, reuse across sessions. `save_study()` writes a self-contained
`.ks` file (rendered rows embedded), and `load_study()` reloads it without the
original ksTFL folder.

```{r}
save_study(study, "my_study.ks")

# Next session
study <- load_study("my_study.ks")
study$tables[["14-3.01"]] <- enrich_context(
  study$tables[["14-3.01"]],
  population = "ITT",
  annotations = list(sap_ref = "Section 9.2")
)
```

## Guardrails

The system prompt instructs the model to never compute or invent statistics:
the numbers in the contexts are final and authoritative, and every referenced
value is cited by output id, population, arm, and row so the writer can verify
it.
