---
title: "Targeted Workflow and Result Chaining"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Targeted Workflow and Result Chaining}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

## Why targeted loading

Clinical studies can produce many outputs, while most writing tasks need only a
small subset. The targeted workflow in ksAI lets you:

- discover available output IDs quickly
- load only the relevant IDs into context
- run skills or free prompts on one or multiple IDs
- save generated analysis in markdown and JSON
- reuse prior analysis in later prompts

## 1. Discover and load only needed IDs

```{r}
library(ksAI)

# Inspect available outputs in a ksTFL meta folder.
ids <- ks_list_ids("path/to/outputs/meta")
ids

# Load only the outputs required for the current task.
study <- ks_load(
  "path/to/outputs/meta",
  ids = c("14-3.01", "14-3.02", "14-7.02")
)
study
```

You can also reload a previously persisted study and optionally filter IDs:

```{r}
study_all <- ks_load("my_study.ks")
study_subset <- ks_load("my_study.ks", ids = c("14-3.01", "14-7.02"))
```

## 2. Run skills on single or multiple IDs

```{r}
chat <- ks_chat(study, model = "qwen3:14b", provider = "ollama")

# Single-ID skill call.
out_describe <- ks_llm(
  chat,
  ids = "14-3.01",
  skill = "describe"
)

# Multi-ID review call (review requires exactly two IDs).
out_review <- ks_llm(
  chat,
  ids = c("14-3.01", "14-3.02"),
  skill = "review"
)

# Skill + extra user instruction. The model follows the language in prompt.
out_spanish <- ks_llm(
  chat,
  ids = "14-7.02",
  skill = "summarize",
  prompt = "Write the final answer in Spanish.",
  audience = "clinician"
)

# Free-form mode (no skill template).
out_free <- ks_llm(
  chat,
  ids = c("14-3.01", "14-7.02"),
  skill = NULL,
  prompt = "Compare efficacy and safety signals and keep the tone regulatory."
)
```

## 3. Save and reload generated outputs

`ks_llm()` returns a `ks_result` object. Save it in both markdown and JSON with
`save_result()`:

```{r}
paths <- save_result(out_spanish, "analysis/out-14-7-02")
paths

# Reload later by base path, .json, or .md path.
out_loaded <- load_result("analysis/out-14-7-02")
out_loaded
```

## 4. Chain prior analysis into a new run

Pass a previous `ks_result` to `prior` so the next run can build on it:

```{r}
out_refined <- ks_llm(
  chat,
  ids = "14-7.02",
  skill = "csr_section",
  title = "Safety Overview",
  prompt = "Keep only statements supported by explicit values.",
  prior = out_loaded
)
```

## 5. Suggested project flow

A practical sequence for writing tasks:

1. `ks_list_ids()` to identify relevant outputs
2. `ks_load(ids = ...)` to create a focused study object
3. `ks_chat()` for model session setup
4. `ks_llm()` for skill-based and free-form generation
5. `save_result()` for traceable deliverables
6. `load_result()` + `prior =` for iterative refinement
