---
title: "Targeted Workflow, Clinical Capsules, and Result Chaining"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Targeted Workflow, Clinical Capsules, and Result Chaining}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

## Overview

ksAI provides two complementary workflows for reasoning over ksTFL outputs.

**Direct context workflow** (Sections 1–4): load a targeted subset of outputs,
render them as markdown or compact text, and run skill-based prompts. This is
the fastest path for single-output tasks and for drafting CSR narratives where
you already know which tables you need.

**Capsule workflow** (Sections 5–8): decompose all outputs into a concept-level
capsule store, enrich capsules with keywords and embeddings, retrieve the
smallest relevant subset for a query, and send only those capsules to a larger
reasoning model. This is the preferred path for exploratory questions across
many outputs and for large studies where injecting full tables would exceed
model context limits.

```
ks_load() → ks_study → ks_context → ks_llm()        # direct
                    ↓
              as_capsules() → ks_capsule_store
              ks_annotate()   (semantic enrichment)
              ks_embed()      (vector embeddings)
              ks_retrieve()   (hybrid retrieval)
              ks_reason()     → ks_result            # capsule
```

---

## 1. Discover and load only needed IDs

```{r}
library(ksAI)

# Inspect available outputs without loading data.
ids <- ks_list_ids("path/to/outputs/meta")
ids

# Load only the outputs required for the current task.
study <- ks_load(
  "path/to/outputs/meta",
  ids = c("14-3.01", "14-3.02", "14-5.01", "14-7.02")
)
study
```

Reload a previously persisted study or filter IDs from an existing `.ks` file:

```{r}
study_all <- ks_load("my_study.ks")
study_subset <- ks_load("my_study.ks", ids = c("14-3.01", "14-7.02"))
```

## 2. Inspect context objects

Each loaded output becomes a `ks_context` — the render-join of specification
and data:

```{r}
ctx <- study[["14-3.01"]]
ctx                       # print: title, columns, span headers, row count
as_markdown(ctx)          # markdown table for reading
as_compact(ctx)           # token-efficient DSL: smaller than markdown
as_json(ctx)              # full JSON (columns + rows)
```

Compact format is ideal for model injection when the table is large. For a
table with span-header groups it emits a one-time legend (`SPANS:`, `COLS:`)
and short keys per row rather than repeating arm labels:

```
TABLE: 14-5.01 | Population: Safety
TITLE: ...
SPANS: A=Placebo (N=86); B=Xanomeline Low Dose (N=84); C=Xanomeline High Dose (N=84)
COLS: n(%), [AEs]
[CARDIAC DISORDERS]
SINUS BRADYCARDIA:  A: 2 (2.3%), [2]  |  B: 7 (8.3%), [10]  |  C: 8 (9.5%), [12]
```

Set the format globally before any LLM call:

```{r}
ks_set_option(context_format = "compact")   # "markdown" | "compact" | "json"
```

## 3. Run skills or free prompts

```{r}
# Preferred: pass study and provider directly (context injected once).
out_describe <- ks_llm(
  study,
  ids = "14-3.01",
  skill = "describe",
  model = "qwen3:14b",
  provider = "ollama"
)

# Multi-ID review (requires exactly two IDs).
out_review <- ks_llm(
  study,
  ids = c("14-3.01", "14-3.02"),
  skill = "review",
  model = "qwen3:14b",
  provider = "ollama"
)

# Extra user instruction on top of a skill template.
out_section <- ks_llm(
  study,
  ids = "14-7.02",
  skill = "csr_section",
  prompt = "Keep only statements supported by explicit values.",
  model = "qwen3:14b",
  provider = "ollama"
)

# Free-form mode (no skill template).
out_free <- ks_llm(
  study,
  ids = c("14-3.01", "14-7.02"),
  skill = NULL,
  prompt = "Compare efficacy and safety signals.",
  model = "qwen3:14b",
  provider = "ollama"
)
```

Alternatively create a persistent chat session — useful when you want to ask
follow-up questions in a conversational thread. Note that `ks_chat()` embeds
all loaded contexts in the system prompt, so set `context_format` before
building the chat and keep `max_rows` within the model's context window:

```{r}
ks_set_option(context_format = "compact", max_rows = 50L)
chat <- ks_chat(study, model = "qwen3:14b", provider = "ollama")
ks_llm(chat, ids = "14-3.01", skill = "describe")
```

## 4. Save, reload, and chain results

`ks_llm()` and `ks_reason()` both return a `ks_result`. Persist with
`save_result()` (writes both `.md` and `.json`) and reload with `load_result()`:

```{r}
save_result(out_section, "analysis/out-14-7-02")
out_loaded <- load_result("analysis/out-14-7-02")
```

Pass a previous result to `prior` so the next run can build on it:

```{r}
out_refined <- ks_llm(
  study,
  ids = "14-7.02",
  skill = "csr_section",
  prompt = "Tighten the narrative; remove any unsupported claims.",
  prior = out_loaded,
  model = "qwen3:14b",
  provider = "ollama"
)
```

---

## 5. Build a Clinical Capsule store

When a study has many outputs or individual tables are too large to fit in one
prompt, decompose them into **clinical capsules** — concept-centric,
hierarchical summaries with parsed statistics, domain tags, and links.

```{r}
# Build capsule store from the entire study (or a single context).
store <- as_capsules(study)
store

# Inspect individual capsules.
names(store$capsules)[1:6]
cap <- store$capsules[["14-5.01::SOC::CARDIAC_DISORDERS"]]
cap                    # print: domain, level, label, row span, child count
cap$stats              # parsed statistics per treatment arm
cap$compact_text       # the text a model will see for this capsule
cap$child_ids          # PT-level capsule IDs nested under this SOC
cap$linked_ids         # listing output IDs in the same study / same domain
```

The hierarchy has three levels by default: OVERALL → SOC → PT (for AE tables),
or OVERALL → PARAM (for continuous endpoints). Parent/child IDs are wired
automatically.

Persist the store alongside your `.ks` file:

```{r}
save_capsules(store, "my_study.ksc")
store <- load_capsules("my_study.ksc")
```

---

## 6. Enrich with keywords and embeddings

### 6a. Pure-R keyword pass (always fast, no model needed)

```{r}
store <- ks_annotate(store)
store$capsules[["14-5.01::SOC::CARDIAC_DISORDERS"]]$keywords
```

This tokenizes `label`, `compact_text`, and existing concepts, strips stop
words, and detects known clinical abbreviations (`TEAE`, `SOC`, `PT`, etc.).

### 6b. Optional small-LLM enrichment pass

A 4B model running locally can identify medical concepts, synonyms, and
richer keywords. LM Studio with `gemma-4-12b` or `qwen3-4b` works well:

```{r}
store <- ks_annotate(
  store,
  model    = "google/gemma-4-12b-qat",
  provider = "lm_studio",
  force    = FALSE   # skip capsules already annotated
)
```

The response is expected to be JSON with keys `concepts`, `synonyms`,
`keywords`. Any non-JSON wrapper text is stripped automatically.

### 6c. Embedding vectors

`ks_embed()` calls the OpenAI-compatible `/v1/embeddings` endpoint —
`text-embedding-nomic-embed-text-v1.5` runs out of the box in LM Studio:

```{r}
ks_set_option(
  embed_model = "text-embedding-nomic-embed-text-v1.5",
  embed_url   = "http://127.0.0.1:1234/v1"
)

store <- ks_embed(store)
length(store$capsules[["14-5.01::SOC::CARDIAC_DISORDERS"]]$embedding)
```

Embeddings are stored in `capsule$embedding` as a numeric vector and
persisted in the `.ksc` file. `ks_retrieve()` falls back to keyword-only
scoring when embeddings are absent.

---

## 7. Retrieve the smallest relevant subset

`ks_retrieve()` scores every capsule against a query using three signals
blended by configurable weights:

| Signal | Default weight | Source |
|--------|---------------|--------|
| Semantic | 0.6 | cosine similarity with query embedding |
| Keyword | 0.3 | token overlap between query and `capsule$keywords` |
| Metadata | 0.1 | exact match on `domain`, `level`, `population` |

```{r}
subset <- ks_retrieve(
  store,
  query  = "cardiac events in the Xanomeline High Dose arm",
  n      = 5L,
  filter = list(domain = "AE", level = "SOC")
)

subset                   # print: query + top-5 ranked capsules with scores
names(subset$capsules)   # selected capsule IDs
subset$scores            # data.frame: capsule_id, semantic, keyword, metadata, score
```

---

## 8. Reason over retrieved capsules

`ks_reason()` combines retrieval and the reasoning step in one call. The
model **never sees the full study** — only the capsule compact texts for the
matched subset, plus its metadata:

```{r}
out <- ks_reason(
  store,
  query    = "Summarize cardiac findings and statistical significance",
  n        = 5L,
  expand   = FALSE,   # TRUE → also include child PT capsules
  model    = "gemma-4-26b-a4b-it-mlx",
  provider = "lm_studio"
)

out                  # ks_result with skill = "reason"
cat(out$response)
```

Set `expand = TRUE` for progressive disclosure: the initial reasoning pass
uses only the retrieved capsules; children are loaded on demand:

```{r}
out_expanded <- ks_reason(
  store,
  query    = "Which PT-level terms drove cardiac disorder incidence?",
  n        = 3L,
  expand   = TRUE,          # pulls in all PT children of matched SOC capsules
  model    = "gemma-4-26b-a4b-it-mlx",
  provider = "lm_studio"
)
```

---

## 9. Suggested project flow

### For focused tasks (1–3 outputs, model context not a constraint)

1. `ks_list_ids()` → identify outputs
2. `ks_load(ids = ...)` → targeted study
3. `ks_set_option(context_format = "compact")` → token-efficient injection
4. `ks_llm()` → generate draft
5. `save_result()` / `load_result()` + `prior =` → iterate

### For broad tasks (many outputs or large tables)

1. `ks_load(ids = NULL)` → full study
2. `as_capsules(study)` → concept-level decomposition
3. `ks_annotate(store)` → keywords (+ optional small-LLM enrichment)
4. `ks_embed(store)` → semantic vectors via local embedding model
5. `save_capsules(store, ...)` → persist for reuse
6. `ks_reason(store, query, n)` → retrieve + reason in one call
7. `save_result()` → traceable deliverable

Capsule stores can be built once from the full study and reused across many
writing sessions without reloading or reprocessing the ksTFL source JSON.
